---
title: "01 Intro to the tidyverse"
author: "Vignesh Venkatachalam"
date: "2021-04"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    collapsed: false
    number_sections: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align="center", message=FALSE,warning=FALSE, cache=TRUE)
```

# Introduction to Tidyverse

Last compiled: `r Sys.Date()`

# Load libraries

Load the necessary libraries at the beginning.

```{r load}
library(tidyverse)

# Excel Files
library(readxl)

```

# Load the data
Load the bike sales data and store into objects.

```{r data, include=FALSE}
bikes_tbl <- read_excel(path = "00_data/01_bike_sales/01_raw_data/bikes.xlsx")
orderlines_tbl <- read_excel(path = "00_data/01_bike_sales/01_raw_data/orderlines.xlsx")
bikeshops_tbl <- read_excel(path = "00_data/01_bike_sales/01_raw_data/bikeshops.xlsx")

```

# Joining the data
```{r join-data}
bike_orderlines_joined_tbl <- orderlines_tbl %>% 
  left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))
```

# Wrangling Data
```{r wrangle-data}
bike_orderlines_wrangled_tbl <- bike_orderlines_joined_tbl %>% 
  
  # 5.1 Separate category name
  separate(col = category,
           into = c("category.1", "category.2", "category.3"),
           sep = "-") %>% 
  
  # 5.2 Add the total price (price * quantity) 
  # Add a column to a tibble that uses a formula-style calculation of other columns
  mutate(total.price = price * quantity) %>% 
  
  # 5.3 Optional: Reorganize. Using select to grab or remove unnecessary columns
  # 5.3.1 by exact column name
  select(-...1, -gender) %>% 
  
  # 5.3.2 by a pattern
  select(-ends_with(".id")) %>% 
  
  # 5.3.3 Actually we need the column "order.id". Let's bind it back to the data
  bind_cols(bike_orderlines_joined_tbl %>% select(order.id)) %>% 
  
  # 5.3.4 You can reorder the data by selecting the columns in your desired order.
  select(order.id, contains("order"), contains("model"), contains("category"),
         price, quantity, total.price,
         everything()) %>% 
  
  # 5.4 Rename columns because we actually wanted underscores instead of the dots
  # (one at the time vs. multiple at once)
  rename(bikeshop = name) %>% 
  set_names(names(.) %>% str_replace_all("\\.", "_"))

```

# Business Insights
## Sales by Year

```{r sales-by-year}
# 6.0 Business Insights ----
# 6.1 Sales by Year ----

library(lubridate)
# Step 1 - Manipulate
sales_by_year_tbl <- bike_orderlines_wrangled_tbl %>% 
  
  # Select columns
  select(order_date, total_price) %>% 
  
  # Add year column
  mutate(year = year(order_date)) %>% 
  
  # Grouping by year and summarizing sales
  group_by(year) %>% 
  summarize(sales = sum(total_price)) %>% 
  
  # Optional: Add a column that turns the numbers into a currency format 
  # (makes it in the plot optically more appealing)
  # mutate(sales_text = scales::dollar(sales)) <- Works for dollar values
  mutate(sales_text = scales::dollar(sales, big.mark = ".",
                                     decimal.mark = ",",
                                     prefix = "",
                                     suffix = " €"))

# sales_by_year_tbl

# Step 2 - Visualize
sales_by_year_tbl %>% 
  
  # Setup canvas with the columns year (x-axis) and sales (y-axis)
  ggplot(aes(x = year, y = sales)) +
  
  # Geometries
  geom_col(fill = "#2DC6D6") + # Use geom_col for a bar plot
  geom_label(aes(label = sales_text)) + # Adding labels to the bars
  geom_smooth(method = "lm", se = FALSE) + # Adding a trendline
  
  # Formatting
  # scale_y_continuous(labels = scales::dollar) + # Change the y-axis. 
  # Again, we have to adjust it for euro values
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".",
                                                     decimal.mark = ",",
                                                     prefix = "",
                                                     suffix = " €")) +
  
  labs(
    title = "Revenue by year",
    subtitle = "Upward Trend",
    x = "", # Override defaults for x and y
    y = "Revenue"
  )

```

## Sales by Year and Category 1
```{r sales-by-year-and-cat}
# 6.2 Sales by Year and Category 1 ----

# Step 1 - Manipulate
sales_by_year_cat_1_tbl <- bike_orderlines_wrangled_tbl %>% 
  
  # Select columns and add a year
  select(order_date, total_price, category_1) %>% 
  mutate(year = year(order_date)) %>% 
  
  # Group by and summarize year and main catgegory
  group_by(year, category_1) %>% 
  summarise(sales = sum(total_price)) %>% 
  ungroup() %>% 
  
  # Format $ Text
  mutate(sales_text = scales::dollar(sales, big.mark = ".",
                                     decimal.mark = ",",
                                     prefix = "",
                                     suffix = " €"))

# sales_by_year_cat_1_tbl

# Step 2 - Visualize
sales_by_year_cat_1_tbl %>% 
  
  # Set up x, y, fill
  ggplot(aes(x = year, y = sales, fill = category_1)) +
  
  # Geometries
  geom_col() + # Run up to here to get a stacked bar plot
  geom_smooth(method = "lm", se = FALSE) +
  
  # Facet
  facet_wrap(~ category_1) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".",
                                                    decimal.mark = ",",
                                                    prefix = "",
                                                    suffix = " €")) +
  labs(
    title = "Revenue by year and main category",
    subtitle = "Each product category has an upward trend",
    fill = "Main category" # Changes the legend name
  )
```

# Challenges
The previous wrangled data is further worked on to separate the location column into city and state and the wrangled data is updated.

## Split location into city and state
```{r update-wrangled-data}
bike_orderlines_wrangled_tbl <- bike_orderlines_wrangled_tbl %>% 
  
  # Separate location into city and state
  separate(col = location,
           into = c("city", "state"),
           sep = ", ")
```

## Sales by State
The steps are the same as the previous ones, year being replaced by state.
```{r sales-by-location}
# 6.3 Sales by Location ----

# Step 1 - Manipulate
sales_by_location_tbl <- bike_orderlines_wrangled_tbl %>% 
  
  # Select columns
  select(state, total_price) %>% 
  
  # Grouping by state and summarizing sales
  group_by(state) %>% 
  summarize(sales = sum(total_price)) %>% 
  
  # Format $ Text
  mutate(sales_text = scales::dollar(sales, big.mark = ".",
                                     decimal.mark = ",",
                                     prefix = "",
                                     suffix = " €"))

# sales_by_location_tbl

# Step 2 - Visualize
sales_by_location_tbl %>%
  
  # Setup columns with thje columns state (x-axis) and sales (y-axis)
  ggplot(aes(x = state, y = sales)) +
  
  # Geometries
  geom_col(fill = "#2DC6D6") +
  geom_label(aes(label = sales_text)) + 
  geom_smooth(method = "lm", se = FALSE) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".",
                                                    decimal.mark = ",",
                                                    prefix = "",
                                                    suffix = " €")) +
  
  labs(
    title = "Revenue by state",
    x = "State",
    y = "Revenue"
  ) +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The last line in R code rotates the values on the x axis of the graph to avoid overlapping while printing out the graph.The state of North Rhine-Westphalia shows the highest revenue among the states.

## Sales by State and Year
The steps are the same as the previous one, category 1 being replaced state.
```{r sales-by-location-and-year}
# 6.4 Sales by Year and State ----

# Step 1 - Manipulate
sales_by_location_year_tbl <- bike_orderlines_wrangled_tbl %>% 
  
  # Select columns and add a year
  select(order_date, total_price, state) %>%
  mutate(year = year(order_date)) %>% 
  
  # Group by and summarize year and state
  group_by(year, state) %>% 
  summarise(sales = sum(total_price)) %>% 
  ungroup() %>% 
  
  # Format $ Text
  mutate(sales_text = scales::dollar(sales, big.mark = ".",
                                     decimal.mark = ",",
                                     prefix = "",
                                     suffix = " €"))

# sales_by_location_year_tbl

# Step 2 - Visualize
sales_by_location_year_tbl %>% 
  
  # Set up x, y and fill
  ggplot(aes(x = year, y = sales, fill = state)) +
  
  # Geometries
  geom_col() + 
  geom_smooth(method = "lm", se = FALSE) +
  
  # Facet
  facet_wrap(~ state) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".",
                                                    decimal.mark = ",",
                                                    prefix = "",
                                                    suffix = " €")) +
  labs(
    title = "Revenue by year and state",
    subtitle = "Most states show an upward trend",
    fill = "State"
  ) +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Saving the Wrangled Data
```{r saving-wrangled-data}
# 7.0 Writing Files ----

# 7.1 Excel ----
library(writexl)
bike_orderlines_wrangled_tbl %>% 
  write_xlsx("00_data/01_bike_sales/02_wrangled_data/bike_orderlines.xlsx")

# 7.2 CSV ----
bike_orderlines_wrangled_tbl %>% 
  write_csv("00_data/01_bike_sales/02_wrangled_data/bike_orderlines.csv")

# 7.3 RDS ----
bike_orderlines_wrangled_tbl %>% 
  write_rds("00_data/01_bike_sales/02_wrangled_data/bike_orderlines.rds")
```