---
title: "02 Data Acqusition"
author: "Vignesh Venkatachalam"
date: "2021-04"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    collapsed: false
    number_sections: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Data Acquisition

Last compiled: `r Sys.Date()`

## Challenge 1: Getting Data via API

I have used the Spotify API to collect information about the Albums of an artist using the artistID as assigned by Spotify. This can be obtained from the Spotify web app on the address bar after you navigate to your favourite artist

```{r load-libraries}
## 1.0 LIBRARIES 

library(tidyverse) 
library(rvest)     
library(xopen)     
library(jsonlite)  
library(glue)      
library(stringi)   
library(httr)
library(knitr)

```

### Client ID for authentication

To obtain a client ID and a secret code, go to the spotify developers console through your spotify account and create an app. Just input those values below in place of "xxxxxxxxxx".

```{r authentication-id}
clientID <- "xxxxxxxxxx" # Add your own clientID here
secret   <- "xxxxxxxxxx" # Add your own client secret code here

# Accessing the authentication token
response <- POST(
  'https://accounts.spotify.com/api/token',
  accept_json(),
  authenticate(clientID, secret),
  body = list(grant_type = 'client_credentials'),
  encode = 'form',
  verbose()
)

mytoken <- content(response)$access_token

HeaderValue <- paste0("Bearer ", mytoken)

```

### Function to Gather Album Data

Just pass the Artist ID of your favorite artist to collect information about their albums on Spotify.

```{r album-list}
get_album_tracks <- function(artistID = "4vC2GtOXDzAfthZ4gnFowC") {
  
  # API request to get the albums of an Artist
  URI <- paste0('https://api.spotify.com/v1/artists/', artistID, '/albums')
  response2 <- GET(url = URI, config = add_headers(Authorization = HeaderValue))
  
  # convert to tibble
  albums <- content(response2)
  
  # Extracting only some information about the albums
  albums_items <- albums$items
  
  # Converting into tibble format
  albums_items_data <- as.data.frame(do.call(cbind, albums_items))
  albums_items_tbl <- as_tibble(nms = names(albums_items_data), t(albums_items_data)) %>% 
    # Removing unnecessary columns
    select(-c(1:8, 11)) %>% 
    set_names(c("Album Name", "Release Date", "Number of Tracks", "Type", "URI"))
}
```

### Example Use case

For this example, I have used the artist of the late Mandolin Maestro Padmashri Mandolin U Srinivas.

```{r example-u-srinivas}
albums_items_tbl <- get_album_tracks(artistID = "4vC2GtOXDzAfthZ4gnFowC")

# Easier to format the look of the table in html file
kable(albums_items_tbl %>% head(n = 10),
      align = "ccccc", 
      caption = "Album information of the Desired Artist")
```

## Challenge 2: Web Scraping

For this challenge, I've decided to use the website of Rose bikes. The necessary libraries were loaded at the start of this document in the previous challenge.

### Function for Scraping for Bike Data

The URL of the specific product category has to passed to this function.

```{r scrape-bike-data}
scrape_bike_data <- function(url) {
  
  html_gravel_bikes <- read_html(url)
  
  # Bike model data
  bike_name <- html_gravel_bikes %>% 
    html_nodes(css = ".catalog-category-model__title") %>% 
    html_text() %>% 
    str_replace_all("[\r\n]", "") %>% 
    enframe(name = "Sl No.", value = "Name")
  
  # Price of the bikes
  bike_price <- html_gravel_bikes %>% 
    html_nodes(css = ".catalog-category-model__price-current-value") %>% 
    html_text() %>% 
    str_replace_all("[\r\n]", "") %>% 
    enframe(name = "Sl No.", value = "Price")
  
  # Combining into one tibble
  bike_info_combined <- left_join(bike_name, bike_price) %>% 
    select(-("Sl No."))
}
```

### Example Use case for the Scraping Function

Here, I have used the Backroad Al model of the Gravel bikes to demonstrate the working of the function. 

```{r scraping-example}
url_home <- "https://www.rosebikes.de/fahrr%C3%A4der/gravel/backroad-al"
bike_info_output <- scrape_bike_data(url_home)

kable(bike_info_output, align = "cc", 
      caption = "Prices of Backroad Al models of Gravel bikes from Rosebikes")
```