---
title: "03 Data Wrangling"
author: "Vignesh Venkatachalam"
date: "2021-05"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    collapsed: false
    number_sections: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Data Wrangling

### Load Libraries
```{r load-libraries}
library(tidyverse)
library(vroom)
library(data.table)
library(lubridate)
library(knitr)
```

### Load Data

I have decided to work on the reduced patent data set. The different tables in the dataset are loaded as shown below with the help of the vroom package.

```{r load-data}

# Patent data table
patent_col_types <- list(
 id         = col_character(),
 date       = col_date("%Y-%m-%d"),
 num_claims = col_double()
)

patent_dt <- vroom(
  file      = "00_data/Patent_data_reduced/patent.tsv",
  delim     = "\t",
  col_types = patent_col_types,
  na        = c("", "NA", "NULL")
)

setDT(patent_dt)

# Patent Assignee data table
patent_assignee_col_types <- list(
  patent_id   = col_character(),
  assignee_id = col_character()
)

patent_assignee_dt <- vroom(
  file      = "00_data/Patent_data_reduced/patent_assignee.tsv",
  delim     = "\t",
  col_types = patent_assignee_col_types,
  na        = c("", "NA", "NULL")
)

setDT(patent_assignee_dt)

# Assignee data table
assignee_col_types <- list(
  id           = col_character(),
  type         = col_character(),
  organization = col_character()
)

assignee_dt <- vroom(
  file      = "00_data/Patent_data_reduced/assignee.tsv",
  delim     = "\t",
  col_types = assignee_col_types,
  na        = c("", "NA", "NULL")
)

setDT(assignee_dt)

# USPC data table
uspc_col_types <- list(
  patent_id = col_character(),
  mainclass_id = col_character(),
  sequence = col_double()
)

uspc_dt <- vroom(
  file      = "00_data/Patent_data_reduced/uspc.tsv",
  delim     = "\t",
  col_types = uspc_col_types,
  na        = c("", "NA", "NULL")
)

setDT(uspc_dt)

```

## Question 1: Patent Dominance

I have decided to employ the data.table approach for solving this challenge. The code readability is reduced as compared to the dplyr approach but this turned out to be faster and also resulted in lesser lines of code.

```{r top-10-us}
# Top 10 US companies using the patent_assignee and the assignee datasets
us_top_10_companies_dt <- merge(x = patent_assignee_dt, y = assignee_dt, by.x = "assignee_id", by.y = "id")[
  type == 2,][!is.na(patent_id) || !is.na(organization),][
    , .(count = .N), by = organization][order(-count), ][1:10,][]

setnames(us_top_10_companies_dt, "organization", "Organization")
setnames(us_top_10_companies_dt, "count", "Total patents")

# Print table in better format to html
kable(us_top_10_companies_dt, align = "cc", caption = "Top 10 US companies by number of patents")
```

## Question 2: Recent Patent Activity

The approach for finding the Top 10 US companies by new patents in August is similar to the previous one

```{r top-10-us-august}
# Top 10 US companies in August uses patent dataset as well
us_top_10_new_august_dt <- merge(x = patent_dt, y = patent_assignee_dt, by.x = "id", by.y = "patent_id")[
  assignee_dt, on = c(assignee_id = "id")
][type == 2,][
  !is.na(id) || !is.na(organization),][, month := month(ymd(date))][month == 8, ][
    , .(count = .N), by = organization
  ][order(-count), ][1:10,][]

setnames(us_top_10_new_august_dt, "organization", "Organization")
setnames(us_top_10_new_august_dt, "count", "Total patents")

kable(us_top_10_new_august_dt, align = "cc", caption = "Top 10 US companies by new patents in August")

```

## Question 3: Innovation in Tech

For this challenge, we need to first get a data.table containing the top 10 companies worldwide. Then we merge this with the uspc, patent_assignee and assignee data.tables to the top 5 USPTO tech main classes.

```{r top-5-USPTO}
# Top 10 companies worldwide
top_10_worldwide_dt <- merge(x = patent_assignee_dt, y = assignee_dt, by.x = "assignee_id", by.y = "id")[
  !is.na(patent_id) || !is.na(organization),][, .(count = .N), by = organization][order(-count),][
    !is.na(organization),][1:10,][]

# Top 5 USPTO tech main classes
top_5_uspto_dt <- merge(x = uspc_dt, y = patent_assignee_dt, by = "patent_id")[
  assignee_dt, on = c(assignee_id = "id")][top_10_worldwide_dt, on = "organization"][
    , .(count = .N), by = mainclass_id][order(-count),][1:5,][]

setnames(top_10_worldwide_dt, "organization", "Organization")
setnames(top_10_worldwide_dt, "count", "Total patents")
kable(top_10_worldwide_dt, align = "cc", caption = "Top 10 companies worldwide by number of patents")

setnames(top_5_uspto_dt, "mainclass_id", "Main class ID")
setnames(top_5_uspto_dt, "count", "Total count")
kable(top_5_uspto_dt, align = "cc", caption = "Top 5 USPTO tech main classes")

```